Data Analysis

### 2023-04-24 ms1_data_analysis.Rmd
RDA axis predictions with phylogeny.

library(ggplot2)
library(dplyr)
library(sjPlot)
library(AICcmodavg)
library(caTools)
library(randomForest)
library(rpart)
library(rpart.plot)
library(C50)
library(printr)
library(betapart)
library(lsmeans)
library(emmeans)
library(multcomp)
library(multcompView)

@@

# ggplot(fidata, aes(x = x, y = y)) +
#   geom_point(size = 2, alpha = 0.4)+
#   stat_smooth(data = fidata %>%
#                 filter(sign == "signif"),
#               method ="lm")+
#   geom_hline(yintercept = 0)+
#   geom_vline(xintercept = 0)+
#   ylab("Effect size of insects")+
#   xlab("Effect size of fungi")+
#   theme_bw()+
#   facet_grid(descriptor~site)

@@

# ENCLOSED IN THE get_ready_for_the_sla_analyses
# Subset a data-set to only control and treatment plots 
# subDat <- dwbRichF %>%
#   filter(treatment %in% subsetVars)
# 
# lgsData <- data.frame()
# 
# for (st in unique(subDat$site)){
#   for (gard in unique(subDat$garden)){
#     
#     stGardDat <- subDat %>%
#             filter(site == st,
#                    garden == gard)
#     
#     # Some quality control
#     if(dim(stGardDat)[1] == 0){
#       next
#       }
#     if(length(unique(stGardDat$treatment)) < 2){
#       next
#       }
#     
#     specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
#     specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
#     
#     spLost <- specInCtr[!(specInCtr %in% specInTrt)]
#     spGained <- specInTrt[!(specInTrt %in% specInCtr)]
#     
#     stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
#                                ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
#     
#     
#     lgsData <- rbind(lgsData, stGardDat)
#     
#     
#   }
# }
```

@@

```{r}
# Calculate position of species and site along a given treatment vector.
function (oridnation, treatment, type = c('site','species')) {
  NULL
}

surfData <- data.woody.brut %>%
  filter(site == "wanang") %>%
  group_by(unified_names) %>%
  summarise(lsla.d = log(mean(sla.d, na.rm=T)),
            water.cont = mean(water.cont, na.rm = T),
            herbivory = mean(perc.herb, na.rm = T))

# Add row names to sla data
surfDataDf <- as.data.frame(surfData)
tibble::column_to_rownames(surfData, var = 'unified_names')
rownames(surfDataDf) <- surfDataDf$unified_names

# Get the treatment vector.
nm <- "f"
# names(ord.list[["w"]])
trtVcts <- ord.list[["w"]]$CCA$biplot[nm, c(1,2)]

# Get species vector.
all.specs <- ord.list[["w"]]$CCA$v[, c(1,2)]
usednms <- surfData$unified_names[surfData$unified_names %in% rownames(all.specs)]
specVcts <- all.specs[usednms,]

@@

testDat <- data.frame(sla = surfDataDf[usednms, ]$lsla.d,
                      water = surfDataDf[usednms, ]$water.cont,
                      herb = surfDataDf[usednms, ]$herbivory,
                      pos = positions[usednms])

ggplot(modDat, aes(x = pos, y = value)) +
  geom_point()+
  stat_smooth(method = "lm") + 
  facet_wrap(~label, scales = 'free')

# Tests + phylogeny!!!!
glsmod1 <- gls(pos ~ sla + water + herb, data = testDat)
glsmod2 <- gls(pos ~ sla +       + herb, data = testDat)
glsmod3 <- gls(pos ~ sla + water       , data = testDat)
glsmod4 <- gls(pos ~               herb, data = testDat)
glsmod5 <- gls(pos ~       water       , data = testDat)
glsmod6 <- gls(pos ~ sla               , data = testDat)

summary(glsmod1)
summary(glsmod2)
summary(glsmod3)
summary(glsmod4)
summary(glsmod5)
summary(glsmod6)

AIC(glsmod1,glsmod2,glsmod3,glsmod4,glsmod5,glsmod6)

summary(glsmod1)

myTree

# Correlate positions with individual species' with their sla. CWM we can leave out because we know it correlates with the .

>>>>

library(AICcmodavg)
library(betapart)
library(caTools)
library(C50)
library(dplyr)
library(emmeans)
library(effects)
library(ggplot2)
library(lsmeans)
library(multcomp)
library(multcompView)
library(printr)
library(randomForest)
library(rpart)
library(rpart.plot)
library(sjPlot)

@@

@@

@@

# Calculate position of species and site along a given treatment vector.

traitPredsCommComp <- function (data = data.woody.brut,
                                ordination = ord.list,
                                phylogeny = myTree.rand.poly,
                                treatment = "f", st = "wanang") {

  surfData <- data %>%
  filter(site == st) %>%
  group_by(unified_names) %>%
  summarise(lsla.d = log(mean(sla.d, na.rm=T)),
            water.cont = mean(water.cont, na.rm = T),
            herbivory = mean(perc.herb, na.rm = T))

  # Add row names to sla data
  surfDataDf <- as.data.frame(surfData)
  rownames(surfDataDf) <- surfDataDf$unified_names

  # Get the treatment vector.
  trtVcts <- ordination[[substr(st, 1, 1)]]$CCA$biplot[treatment, c(1,2)]

  # Get species vector.
  all.specs <- ordination[[substr(st, 1, 1)]]$CCA$v[, c(1,2)]
  usednms <- surfDataDf$unified_names[surfDataDf$unified_names %in% rownames(all.specs)]
  specVcts <- all.specs[usednms,]

@@

  testDat <- data.frame(sla = surfDataDf[usednms, ]$lsla.d,
                        water = surfDataDf[usednms, ]$water.cont,
                        herb = surfDataDf[usednms, ]$herbivory,
                        pos = positions[usednms],
                        unified_names = usednms)

  # print(ggplot(modDat, aes(x = pos, y = value)) +
  #   geom_point()+
  #   stat_smooth(method = "lm") + 
  #   facet_wrap(~label, scales = 'free'))

  # Phylogeny ready
  testDatTree <- testDat %>%
    filter(unified_names %in% phylogeny$tip.label)

  # Tests + phylogeny!!!!
  glsmod1 <- gls(pos ~ sla + water + herb,
                 data = testDatTree,
                 method = "ML",
                 correlation = corMartins(2.718282,
                       phy = phylogeny,
                       form = ~unified_names,
                       fixed = T),
                 na.action = na.omit)

  # glsmod2 <- gls(pos ~ sla +       + herb, 
  #                data = testDatTree,
  #                method = "ML",
  #                correlation = corMartins(2.718282,
  #                      phy = myTree.rand.poly,
  #                      form = ~unified_names,
  #                      fixed = T),
  #                na.action = na.omit)
  # 
  # glsmod3 <- gls(pos ~ sla + water       ,
  #                data = testDatTree,
  #                method = "ML",
  #                correlation = corMartins(2.718282,
  #                      phy = myTree.rand.poly,
  #                      form = ~unified_names,
  #                      fixed = T),
  #                na.action = na.omit)
  # 
  # glsmod4 <- gls(pos ~               herb,
  #                data = testDatTree,
  #                method = "ML",
  #                correlation = corMartins(2.718282,
  #                      phy = myTree.rand.poly,
  #                      form = ~unified_names,
  #                      fixed = T),
  #                na.action = na.omit)
  # 
  # glsmod5 <- gls(pos ~       water       ,
  #                data = testDatTree,
  #                method = "ML",
  #                correlation = corMartins(2.718282,
  #                      phy = myTree.rand.poly,
  #                      form = ~unified_names,
  #                      fixed = T),
  #                na.action = na.omit)
  # glsmod6 <- gls(pos ~ sla               ,
  #                data = testDatTree,
  #                method = "ML",
  #                correlation = corMartins(2.718282,
  #                      phy = myTree.rand.poly,
  #                      form = ~unified_names,
  #                      fixed = T),
  #                na.action = na.omit)

  # Summary

  tab_model(glsmod1, title = paste(st, treatment))

  # Prediction of the full model
  # effx1 <- effect("sla", glsmod1, partial.residuals=T)
  # effx2 <- effect("water", glsmod1, partial.residuals=T)
  # effx3 <- effect("herb", glsmod1, partial.residuals=T)

  # plot(effx1)
  # plot(effx2)
  # plot(effx3)

traitPredsCommComp(treatment = "f", st = "wanang")
traitPredsCommComp(treatment = "f", st = "numba")
traitPredsCommComp(treatment = "f", st = "yawan")

traitPredsCommComp(treatment = "i", st = "wanang")
traitPredsCommComp(treatment = "i", st = "numba")
traitPredsCommComp(treatment = "i", st = "yawan")
